\typeout{}
\typeout{Compiling references.tex}
\newcommand{\icgafour}{
	{\em Proceedings of the Fourth International Conference
	on Genetic Algorithms,} }
\newcommand{\illigal}[1]
	{(IlliGAL Report No. #1).
	Urbana: University of Illinois, Illinois Genetic Algorithms Laboratory.
	}
\newcommand{\foga}{
	{\em Foundations of Genetic Algorithms,}
}

\begin{thereferences}
\bibitem{walsh} Beauchamp, K. G. (1975). {\em Walsh functions and their 
	applications}, Academic Press.
\bibitem{funopt} Bethke, A. D. (1981), {\em Genetic algorithms as function 
	optimizers} (Doctoral dissertation, University of Michigan).
	{\em Dissertation Abstracts International, 41(9),} 3503B.
	(University Microfilms No. 8106101).
\bibitem{bioconcept} Bledsoe, W. W. (1961). {\em The use of biological 
	concepts in the
	analytical study of systems}. Paper presented at the ORSA-TIMS
	National Meeting, San Francisco, CA.
\bibitem{nonuniform} Bridges, C., Goldberg, D. E. (1989).
	{\em A note on the nonuniform Walsh-schema transform.}
	TCGA Report No. 89004.
	Tuscaloosa, AL: University of Alabama, Department of Engineering
	Mechanics, The Clearinghouse for Genetic Algorithms.
\bibitem{challenge} Das, R., \& Whitley, D. (1991).
	The only challenging problems are deceptive:
	global search by solving order-1 hyperplanes.
	\icgafour 166-173.
\bibitem{davis} Davis, L. (1991).
	Bit-climbing, representational bias, and test suite design.
	\icgafour 18-23.
\bibitem{traps} Deb, K. \& Goldberg, D. E. (1991).
	{\em Analyzing Deception in Trap Functions.}
	\illigal{91009}
\bibitem{sufficient} Deb, K., \& Goldberg, D. E. (1992).
	{\em Sufficient conditions for
	deceptive and easy binary functions.} 
	\illigal{92001}
\bibitem{epistasis} Davidor, Y. (1991).
	Epistasis variance: a viewpoint on GA-hardness.
	\foga 23-35.
\bibitem{forrest} Forrest, S., \& Mitchell, M. (1991).
	The performance of genetic algorithms on Walsh polynomials:
	some anomalous results and their explanation.
	\icgafour 182-189.
\bibitem{mindecp} Goldberg, D. E. (1987). Simple genetic algorithms and the
	minimal, deceptive problem. {\em Genetic Algorithms and Simulated
	Annealing}, 74-88.  Los Altos: Morgan Kaufmann Publishers, Inc.
\bibitem{bible} Goldberg, D. E. (1989a). {\em Genetic algorithms in search,
	optimization, and machine learning}. Reading, MA: Addison-Wesley.
\bibitem{goldberg89b} Goldberg, D. E. (1989b).  Genetic algorithms and
	Walsh functions: Part I, a gentle introduction. {\em Complex
	Systems, 3,} 129-152.
\bibitem{goldberg89c} Goldberg, D. E. (1989c).  Genetic algorithms and
	Walsh functions: Part II, deception and its analysis. {\em Complex
	Systems, 3,} 153-171.
\bibitem{real} Goldberg, D. E. (1990a). Real-coded genetic algorithms,
	virtual alphabets, and blocking.
	{\em Complex Systems, 5}(2), 139-168.
\bibitem{goldberg90b} Goldberg, D. E. (1990b).  {\em Construction of
	higher-order deceptive functions using low-order Walsh coefficients.}
	\illigal{90002}
\bibitem{clark} Goldberg, D. E., Deb, K., \& Clark, J. H. (1991).
	{\em Genetic Algorithms, noise, and the sizing of populations.}
	\illigal{91010}
\bibitem{multimodal} Goldberg, D. E., Deb, K., \& Horn, J. (1992). {\em Massive
	multimodality, deception, and genetic algorithms.}
	\illigal{92005}
\bibitem{variance} Goldberg, D. E., \& Rudnick, M. (1991). Genetic algorithms
	and the variance of fitness. {\em Complex Systems, 5,} 265-278.
\bibitem{sbbh} Grefenstette, J. J. (1991). Building block hypothesis
	considered harmful. {\em Genetic Algorithms Digest} [email
	journal], {\em 5}(19).
\bibitem{NP} Hart, W. \& Belew, R.
	Optimizing an arbitrary function is hard for the genetic
	algorithm.
	\icgafour 190-195.
\bibitem{holland} Holland, J. H. (1975).  {\em Adaption in natural and
	artificial systems.}  Ann Arbor, MI: University of Michigan Press.
\bibitem{hadamard} Homaifar, A., Qi, X., \& Fost, J. (1991). Analysis and
	design of a general GA deceptive problem,
	\icgafour 196-203.
\bibitem{hadrev} Homaifar, A., \& Qi, X. (1990).
	Analysis of GAs deception by Hadamard transform.
	IASTED International Symposium Machine Learning and Neural
	Networks.  New York.
\bibitem{order} Kargupta, H., Deb, K., \& Goldberg, D. E. (1992).
	{\em Ordering genetic algorithms and deception.}
	\illigal{92006}
\bibitem{rugged} Kauffman, S. A. (1989).
	Adaption on rugged fitness landscapes.
	In D.~L.~Stein, ed., {\em Lectures in the sciences of
	complexity,} pp. 527-618.  Reading: Addison-Wesley.
\bibitem{physics} Kauffman, S. A. (1990).
	Requirements for evolvability in complex systems:
	orderly dynamics and frozen components. {\em Physica D, 42,}
	135-152.
\bibitem{kaufleg} Kauffman, S. \& Levin, S. (1987).
	Towards a general theory of adaptive walks on rugged landscapes.
	{\em Journal of theoretical Biology 128,} 11-45.
\bibitem{simanneal} Kirkpatrick, S. , Gelatt, C., \& Vecci, M.  (1983).  
	Optimization by simulated annealing. {\em Science, 220}, 671-680.
\bibitem{lv90a} Liepins, G. E., \& Vose, M. D. (1990a).
	Polynomials, basic sets, and deceptiveness in genetic algorithms.
	{\em Complex Systems, 5,} 45-61.
\bibitem{lv90b} Liepins, G. E., \& Vose, M. D. (1990b).
	Representational issues in genetic algorithms.
	{\em Journal of Experimental and Theoretical Artificial Intelligence,
	2,} 101-115.
\bibitem{liepins} Liepins, G. E., \& Vose, M. D. (1991). Deceptiveness and 
	genetic algorithm dynamics. \foga 36-50.
\bibitem{CA} Lipsitch, M. (1991).
	Adaption on rugged landscapes generated by iterated local
	interactions of neighboring genes.
	\icgafour 128-135.
\bibitem{fourfall} Lighthill, M. J. (1959).
	{\em Introduction to Fourier analysis and generalised functions.}
	New York: Cambridge University Press.
\bibitem{manderick} Manderick, B., de Weger, M., \& Spiessens, P. (1991).
	The genetic algorithm and the structure of the fitness landscape.
	\icgafour 143-150.
\bibitem{mason} Mason, A. J. (1991). Partition coefficients, static deception
	and deceptive problems for non-binary alphabets.
	\icgafour 210-214.
\bibitem{whatsdecep} Mitchell, M. \& Forrest, S. (1991).
	{\em What is deception anyway?  And what does it have to do with
	GAs?}
	Unpublished manuscript.
\bibitem{royalroad} Mitchell, M. Forrest, S., \& Holland, J. (1991).
	{\em The royal road for genetic algorithms: fitness
	landscapes and GA performance.}
	Unpublished manuscript.
\bibitem{forma} Radcliffe, N. (1991).
	Forma analysis and random respectful mutations.
	\icgafour 222-229.
\bibitem{ruddy} Rudnick, M. (1991). {\em Genetic algorithms and the variance
	of fitness.}  Unpublished Doctoral Dissertation, Oregon
	Graduate Institute, Department of Computer Science and
	Engineering, Beaverton, OR.
\bibitem{signal} Rudnick, M., \& Goldberg, D. E. (1991). {\em Signal, noise, and
	genetic algorithms.}
	\illigal{91005}
\bibitem{spurrious} Schaffer, J. D., Eshelman, L. J., \&
	Offutt, D. (1991).
	Spurious correlations and premature convergence in genetic
	algorithms.
	\foga 102-112.
\bibitem{weird} Sikora, R. (1991). {\em Analysis of deception for
	permutation problems.}  Unpublished manuscript.
\bibitem{fsa} Szu, H., \& Hartley, R. (1987). Fast simulated annealing.
	{\em Physics Letters A, 122} (3,4), 157.
\bibitem{tanese} Tanese, R. (1989).
	Distributed genetic algorithms for function optimization.
	(Doctoral Dissertation, University of Michigan).
	{\em Dissertation Abstracts International, 50,} 5180B.
	(University Microfilms No. 90-01722).
, Ann Arbor, MI).
\bibitem{fourier} Tolstov, G. P. (1962).
	{\em Fourier series.}
	Englewood Cliffs, NJ: Prentice-Hall, Inc.
\bibitem{vose} Vose, M., \& Liepins, G. (1991). Schema disruption.
	\icgafour 237-242.
\bibitem{wein87} Weinberger, E. (1987).  A stochastic generalization
	of Eigen's model of natural selection (Doctoral dissertation,
	New York University).  {\em Dissertation Abstracts International,
	48,} 2000B. (University Microfilms No. 87-22798)
\bibitem{wein88} Weinberger, E. (1988).  A more rigorous derivation of some
	properties of uncorrelated fitness landscapes [Letters to the
	editor].  {\em Journal of Theoretical Biology, 134}, 125-129.
\bibitem{weinberger} Weinberger, E. (1990). Correlated and uncorrelated
	fitness landscapes and how to tell the difference.
	{\em Biological Cybernetics, 63,} 325-336.
\bibitem{whitley91} Whitley, D. (1991a). Fundamental principles of
	deception in genetic search.  
	\foga 221-241.
\bibitem{whitley91b} Whitley, D. (1991b).  Deception, dominance and
	implicit parallelism (Technical Report No. CS-91-120).
	Fort Collins: Colorado State University, Department of
	Computer Science.
\bibitem{wilson91} Wilson, S. (1991). GA-easy does not imply
	steepest-ascent optimizable.
	\icgafour 85-89.
\bibitem{wright} Wright, A. H. (1991).
	Genetic algorithms for real parameter optimization.
	\foga 205-218.
\end{thereferences}


